HGP analysis

### 2019-03-02 rick/1KGP.Rmd
update

```{r}
bas_dt<- read_tsv( "average_quality_chrom21.txt")
```

## Plot

@@

# Load Gentype data

Load VCF file into a sparse matrix

```{r}
p <- pipe("zcat ALL.chr20.phase3_shapeit2_mvncall_integrated_v5a.20130502.genotypes.vcf.gz  |
          sed /^#/d  |
          cut  -f '10-' |
          ./a.out | 
          cut -f '1-2'")
x <- read.table(p,
                colClasses=c("integer","integer"), 
                fill=TRUE,
                row.names=NULL)
```

```{r}
# Convert to a sparse matrix of people (rows) x variant (columns)
chr20 <- sparseMatrix(i=x[,2], j=x[,1], x=1L)

# Inspect the dimensions of this matrix
print(dim(chr20))
```

```{r}
rm(x)
gc()
```

@@

@@

```{r}
chr20_sub <- chr20[,1:100000]
```

@@

```{r}
chr20_mtx <- t(as.matrix(chr20[,1:100000]))
```

compute the first three factors including the intercept

```{r}
LF <- lfa(chr20_mtx, 4)
dim(LF)
```

computes the individual-specific allele frequencies

```{r}
chr20sub_ial <- af(chr20_mtx, LF)

fit_model_1 <- function(data, qual){

  fit <- glm(genotype ~ qual + offset(iaf), family ="binomial", data=data)
  return(fit)

plot_data <- function(data,qual){

      p <- ggplot(data,
       aes(qual, genotype, color=iaf)) +
      geom_point() +
      theme_bw()
  return(p)

}
```

loci <- c(1:1000)
names(loci) <- loci

data_list <- future_map(loci, 
                 to_datalist,
                 data_genotype=chr20_mtx,
                 data_iaf=chr20sub_ial)

```{r}
data_list <- 
  enframe(data_list,
          name = "locus", 
          value = "dataset")
```

```{r}
avg_qual_mean <- bas_dt$avg_qual_mean

data_list <- 
  data_list %>%
  mutate(
    fit = future_map(dataset,
                     ~ glm(genotype ~ avg_qual_mean + offset(iaf),
                           family ="binomial",
                           data=.x)),
    tidied = future_map(fit, tidy),
    glanced = future_map(fit, glance)
)
```

get all p-values

```{r}
target_loci <- c("4", "5")
```

>>>>

 Plot

@@

```{r}
bas_dt <- read_tsv( "average_quality_chrom21.txt")
avg_qual_mean <- bas_dt$avg_qual_mean
```

get_data <- function(loci){

  p <- pipe("zcat ALL.chr20.phase3_shapeit2_mvncall_integrated_v5a.20130502.genotypes.vcf.gz  |
          sed /^#/d  |
          cut  -f '10-' |
          ./a.out | 
          cut -f '1-2'")
  x <- read.table(p,
                colClasses=c("integer","integer"), 
                fill=TRUE,
                row.names=NULL)

  data <- sparseMatrix(i=x[,2], j=x[,1], x=1L)

  data <- t(as.matrix(data[,loci]))

  return(data)

get_af_data <- function(data, d=4){

  # compute the first d factors including the intercept
  LF <- lfa(data, d)
  # compute the individual-specific allele frequencies
  data_iaf <- af(data, LF)
  loci <- c(1:dim(data)[1])
  names(loci) <- loci

  data_list <- future_map(loci, 
                 to_datalist,
                 data_genotype=data,
                 data_iaf=data_iaf)

  data_list <- 
  enframe(data_list,
          name = "locus", 
          value = "dataset")

  return(data_list)

get_predicted_mean <- function(model_fit) {
 predicted <- 
   augment(model_fit, type.predict = "response") %>%
   select("genotype","offset.iaf.",".fitted") %>%
   rename(iaf="offset.iaf.", mu=".fitted" )
  return(predicted)
}

get_stats <- function(model_fit){
  stats_dt <- 
    tidy(model_fit)  %>%
    filter(term=="avg_qual_mean") %>%
    select(estimate, p.value) %>%
    mutate(deviance= anova(model_fit)["Deviance"][[1]][2])
  return(stats_dt)
}

fit_model <- function(dataset){

   fit <- glm(genotype ~ avg_qual_mean + offset(iaf),
                           family ="binomial",
                           data=dataset)

   stats <- get_stats(fit)
   predicted <- get_predicted_mean(fit)

   return(list(stats, predicted))
}

get_plot <- function(data){

 p <- ggplot(bind_cols(data, 
                   bas_dt %>% 
                     select(Population,
                            avg_qual_mean))) +
    geom_point(aes(avg_qual_mean, 
                   genotype),
               size=.5,
               alpha= 0.4) +
    geom_line(aes(avg_qual_mean,
                  iaf, 
                  colour="af"),
              linetype="dashed") +
    geom_line(aes(avg_qual_mean,
                  mu,
                  colour="mu")) +
      theme_bw() +
    facet_wrap(~Population)

  return(p)
}
```

```{r}
run_workflow <- function(loci){

  data <- get_data(loci)
  data_list <- get_af_data(data)

  rm(data)
  gc()

  data_list <- 
    data_list %>%
  #slice(1:10) %>%
  transmute(
    locus=locus,
    fit1 = future_map(dataset, fit_model))  %>%
  transmute(
    locus=locus,
    stats = future_map(fit1, c(1)),
    predicted =future_map(fit1, c(2))) %>%
    unnest(stats) %>%
    mutate(plotted= future_map(predicted, get_plot)) %>%
    arrange(-deviance)

  return(data_list)

```{r}
data_list <- run_workflow (loci=c(1:100))
data_list
```

```{r}
n <-5
target_loci <- 
  data_list %>% 
  slice(1:n) %>% 
  pull(locus)
target_loci 
```

```{r}
data_list_subset %>%
  filter(locus %in% target_loci )%>%
  pull(plotted)
```

## Draft

@@

data_list
```

@@

```{r}
chr20_sub <- chr20[,1:50000]
```

@@
